---
author: |
  | William Johansson Joaquín Barrutia
title: "Kicking the Odds: A Bayesian Framework for Football Match Outcome Prediction in Allsvenskan"
institute: Uppsala University
date: \today
output: 
  binb::metropolis:
    citation_package: natbib
    includes:
      in_header: miscelanea/header.tex
urlcolor: blue
bibliography: miscelanea/main.bib
biblio-style: abbrvnat
beameroption: "show notes"
toc: true
vignette: >
  %\VignetteIndexEntry{binb Metropolis Demo}
  %\VignetteKeywords{binb,vignette}
  %\VignettePackage{binb}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r,setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)

library(knitr)    
library(ggplot2)
library(png) 
library(rnaturalearth)
library(gridExtra)
library(grid)
library(jpeg)
```

# Introduction

## Motivation {#id1.1}

-   On 2022-10-03, the Swedish betting site SvenskaSpel introduced a new betting game called \textbf{\alert{Fullträff.}}

-   No one has ever been able to predict all 13 games.

- Can a Bayesian Framework help in predicting Allsvenskan results?

## Literature Review {#id1.2}

- Statistics has extensive literature dedicated to forecasting soccer match outcomes identifying profitable betting opportunities.

-   Early works that gained popularity employed:

    - Poisson models to predict soccer scores [@maher_modelling_1982; @dixon_modelling_1997].

- Authors have recently moved to address this prediction problem within a \alert{Bayesian framework} [@baio_bayesian_2010; @robberechts_bayesian_2021]

- Our paper leverages the fantastic tools developed by @egidi_fitting_2022 to:

    - Predict score differences before the commencement of matches within a Bayesian framework.
    
    - We focus on an under-explored league, such as the \alert{Allsvenskan}.

## Data {#id1.3}

-   We constructed a dataset using information from two distinct sources:

    1. [football-data.co.uk](https://www.football-data.co.uk/) website, which has updated match results for leagues around the world.
    
    2. a [https://www.kaggle.com/](https://www.kaggle.com/datasets/stefanoleone992/ea-sports-fc-24-complete-player-dataset) dataset with ratings for every team featured in the EA Sports FIFA 22 Video Game.
    
- Our final dataset includes each of the games of the 16 teams in Allsvenskan in 2022 Season ($240 = 2{16 \choose 2}$)

- For each match we have the following variables:

    - full-time goals scored by both the home and away teams 
    - match date
    - overall average of the FIFA 22 teams' rating, computed using attack, midfield and defense ratings

# Models

- We used two different kinds of models \alert{Static} and \alert{Dynamic}

- The proposed models assume that the data followed two distributions: \alert{Bivariate Poisson} and \alert{Skellam}.

- We will describe the easiest base model in detail which is the Static \alert{Bivariate Poisson}

    - it is really easy to modify the base model to have a Skellam distribution or to change it from Static to Dynamic.

## Static Bivariate Poisson

The model is described as

```{=latex}
\begin{equation} \label{eq:1}
    y_n^h, y_n^a|\lambda_{1n}, \lambda_{2n}, \lambda_{3n} \sim BivPoisson(\lambda_{1n}, \lambda_{2n}, \lambda_{3n})
\end{equation}
\begin{equation} \label{eq:1.1}
  log(\lambda_{1n}) = \mu + home + att_{h_n}+def_{a_n} + \frac{\gamma}{2}(rank_{h_n} - rank_{a_n})
\end{equation}
\begin{equation} \label{eq:1.2}
  log(\lambda_{2n}) = \mu + att_{a_n}+def_{h_n} - \frac{\gamma}{2}(rank_{h_n} - rank_{a_n})
\end{equation}
\begin{equation} \label{eq:1.3}
    log(\lambda_{3n})=\beta_0+ \gamma_1 \beta_{h_n}+\gamma_2 \beta_{a_n}+\gamma_3 \beta w_n
\end{equation}
```

where $(y_n^h, y_n^a)$ denote the number of goals scored by home and away team in $n$-th game,  $\lambda_{1n},\lambda_{2n}$ represent home and away teams' scoring rates. Each $\lambda_{in}$, $\mu$ represent a constant intercept, $home$ is the benefit from playing at your stadium, $att_{h_n}$ and $def_{h_n}$ represent the attack and defense ability of home team $h$ playing in game $n$. ($rank_{h_n} - rank_{a_n}$) are FIFA ranking incorporated as a predictor.

## Static Bivariate Poisson

The priors for the parameters of each team $t$ in the model, are:

```{=latex}
\begin{equation} \label{eq:1.4}
    att_t\sim N(\mu_{att}, \sigma_{att})
\end{equation}
\begin{equation} \label{eq:1.5}
    def_t\sim N(\mu_{def},\sigma_{def})
\end{equation}
\begin{equation} \label{eq:1.6}
    \sigma_{att},\sigma_{def}\sim Cauchy^+(0, 5)
\end{equation}
\begin{equation} 
\begin{split} \label{eq:1.6}
    \gamma \sim N(0,1)
\end{split}
\end{equation}
```

For our sensitivity analysis we changed these priors.

# Results

## Model Comparisons

- We approximated the posterior distributions for the parameters of the models employing the Markov Chain Monte Carlo (MCMC).

- Pointwise out-of-sample prediction accuracy
```{=latex}
\begin{table}[!htbp]
\label{table2}
\vskip 0.15in
\begin{center}
\begin{tiny}
\begin{sc}
\begin{tabular}{lccccr}
\toprule
Model & Priors & $\text{elpd}_{loo}$ & $p_{eff}$ & looic\\
\midrule
BivPoiss Dyn & original & -597.39 & 17.75 & 1,194.79\\
& & (11.75) & (1.39) & (23.51)\\
BivPoiss Stat & original &-596.14 & 19.61 & 1,192.28\\
& & (11.29) & (1.35) & (22.59)\\
Skellam Dyn & original & -399.5 & 25 & 799.01\\
& & (11.06) & (2.54) & (22.12)\\
\alert{Skellam Stat} & \alert{original} & \alert{-395.86} & \alert{17.61} & \alert{791.72}\\
& & \alert{(11.01)} & \alert{(1.83)} & \alert{(22.03)}\\
BivPoiss Dyn & modified &-596.57 & 19.98 & 1,193.13\\
& & (11.58) & (1.53) & (23.17)\\
BivPoiss Stat & modified & -596.26 & 19.19 & 1,192.53\\
& & (11.36) & (1.33) & (22.73)\\
Skellam Dyn & modified & -399.67 & 26.18 & 799.33\\
& & (10.98) & (2.81) & (21.96)\\
Skellam Stat & modified & -395.99 & 17.59 & 791.98\\
& & (11.10) & (1.83) & (22.196)\\
\bottomrule
\end{tabular}
\end{sc}
\end{tiny}
\end{center}
\vskip -0.1in
\end{table}
```

## Model Checking

- The score difference frequencies seem to be appropriately captured by the Static Skellam model’s replications $(y^h-y^a)^{rep}$  (Figure 1).

- Figure 2: overlap between observed goal difference density and replicated densities,the Static Skellam model reasonably captures the goal difference.

```{r, message = FALSE, echo=FALSE, fig.align="center", out.width = "80%"}
library(cowplot)
library(ggplot2)
library(magick)

img_path <- "../data/plots/pp_foot Static Skellam Original Priors.jpeg"
img_path_2 <- "../data/plots/PPC density Static Skellam Original Priors.jpeg"

p1 <- ggdraw() + draw_image(img_path, scale = 0.9)
p2 <- ggdraw() + draw_image(img_path_2, scale = 0.9)

plot_grid(p1, p2)
```

## Predictive Accuracy 

We can compute the probabilities of the posterior results for all matches used in test set --last 2 weeks of 22 Season $\widetilde{(y^h-y^a)}$. The red square represents the observed result.

```{r, message = FALSE, echo=FALSE, fig.align="center", out.width = "50%"}
img_path <- "../data/plots/Posterior match probabilities Static Skellam Original Priors.jpeg"
include_graphics(img_path)
```

## Predictive Accuracy 

We can also show the computed probabilities for a home win for the 32 test matches. The red cells denote more likely home wins.

```{r, message = FALSE, echo=FALSE, fig.align="center", out.width = "56%"}
img_path <- "../data/plots/Round_robin Skellam Original Priors.jpeg"
include_graphics(img_path)
```

## Predictive Accuracy 

Finally, we can reconstruct a final rank league table, predicting position and total amount of points at the end of the season using the in-sample replications $(y^h-y^a)^{rep}$ to compute credible intervals.

```{r, message = FALSE, echo=FALSE, fig.align="center", out.width = "48%"}
img_path <- "../data/plots/foot_rank Skellam Original Priors.jpeg"
include_graphics(img_path)
```

# Conclusions

- The model \alert{accurately} predicted \alert{18} out of 32 games.

- Simulating the simplest betting strategy

    - betting 1 unit on the most probable outcome according to the model in each of the 32 test-games.
    
- The profit of betting 100 SEK using this strategy would be \alert{569 SEK}.

- The model can be developed even further, with more advanced measures of team rankings or abilities

- The betting strategy can also be improved.
